Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-06-19 forecast/paper/paper.tex
Full pass of edits

% Please add a significance statement to explain the relevance of your work
\significancestatement{Forecasting is a critical element of the public health
  response to any fast-moving epidemic or pandemic.  Predicting the future
  spread of some disease process, based on disease activity in the recent
  past---say, using what is known as autoregressive modeling in the time series
  literature---is a basic and widely-used method.  In this paper we study
  whether including auxiliary indicators as additional features in such an
  autoregressive model can be useful in COVID-19 case forecasting and hotspot
  prediction.  We find that indicators based on de-identified medical insurance 
  claims, self-reported symptoms through online surveys, and COVID-related
  internet search activity all provide a nontrivial boost in accuracy, though
  their utility generally depends the pandemic's local dynamics (upswing,
  downswing, or flat period) at prediction time.
}

@@

We consider prediction of future COVID-19 case rates or case hotspots (to be
defined precisely shortly).  By case rate, we mean the case count per 100,000
people (the standard in epidemiology).  We use reported case data aggregated by 
JHU CSSE \cite{Dong:2020}, which, like the auxiliary indicators that we use to supplement the
basic purely autoregressive models, is accessible through the COVIDcast API
\cite{CovidcastAPI}.

The indicators we focus on provide information not generally available from
standard public health reporting. Among the many auxiliary indicators collected
in the API, we study the following five:  
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Trends and Impact Survey COVID-like illness in the community
  (CTIS-CLI-in-community): The estimated percentage of the population who know
  someone in their local community that is sick, based on Delphi's surveys of
  Facebook users, called the COVID Trends and Impact Survey. 
\item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
  computed based on de-identified medical insurance claims from other health
  systems partners.  
\item Google Search Trends for anosmia and ageusia (Google-AA): A measure of 
  Google search volume for queries relating to anosmia or
  ageusia (loss of smell or taste), based on Google's Search Trends data set. 
\end{itemize}
In short, we choose these indicators because, conceptually speaking, they
measure aspects of an individual's disease progression that would plausibly
precede the occurence of (at worst, co-occur  with) the report of a positive
COVID-19 test, through standard public health reporting streams.

@@

HRRs, by their definition, would be most relevant for forecasting hospital
demand.  We have chosen to focus on cases (forecasting and predicting
hotspots) at the HRR level because the indicators considered should be more 
useful in predicting case activity rather than hospital demand. The \attn{former?} is
intuitively more contemporaneous to the events that are measured by
the given five indicators. Predicting case rates (and hotspots) at the HRR level
is still a reasonable goal in its own right; and moreover, it could be used to
feed predicted case information into downstream hospitalization models.

@@

In this paper, all models are fit with ``vintage'' training data. This means
that for a given prediction date, say, September 28, 2020, we train models 
using only the data that would have been available to us \textit{as of}
September 28 (imagine that we can ``rewind'' the clock to September 28 and query
the COVIDcast API to get the latest data it would have had available at that
point in time.)  This is made possible by the COVIDcast API's comprehensive data   
versioning system (described in more detail in our companion paper).  We also
use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the
process of training arbitrary prediction models over a sequence of prediction
dates, by constructing the proper sequence of vintage training data sets.

\begin{figure}[tb!]
  \includegraphics[width=\columnwidth]{fig/revisions.pdf}
  \caption{This figure displays revision behavior for two signals in the HRR
    containing New York 
    City.  Each colored line corresponds to the data as reported on a particular 
    date (\textit{as of} dates varying from September 28 through October
    19). The left panel shows the DV-CLI signal, which was regularly revised
    throughout the period, though the effects fade as we look further back in   
    time. In contrast, the right panel shows case rates reported by JHU CSSE,
    which remain ``as reported'' on September 28, with a huge spike towards the 
    end of this period, until a major correction is made on October 19, which
    brings down this spike and affects all prior data as well.}
  \label{fig:vintage}
\end{figure}

Vintage training data means different things, in practice, for different
signals. The three signals based on medical claims, CHNG-CLI, CHNG-COVID, and
DV-CLI, are most often 3-5 days latent, and subject to a considerable but
regular degree of revision or ``backfill'' after their initial publication date.
The survey-based signal, CTIS-CLI-in-community, is 2 days latent, and rarely
undergoes any revision at all.  The target variable itself, reported COVID-19
case rates, is 1 day latent, and exhibits frequent, unpredictable revisions
after initial publication.  Compared to the revision
pattern in the medical claims signals, which are much more systematic in nature,
revisions in case reports can be highly erratic. Big spikes or other anomalies
can occur in the data as reporting backlogs are cleared, changes in case
definitions are made, etc. Groups like JHU CSSE then work tirelessly to correct
such anomalies after first publication (e.g., they will attempt to
back-distribute a spike when a reporting backlog is cleared, by working with a
local authority to figure out how this should best be done), which can result
in very nontrivial revisions.  See
Figure~\ref{fig:vintage} for an example.

@@

Why do we define the response variables via 7-day averaging? The short answer 
is robustness: averaging stabilizes the case time series, and accounts for 
uninteresting artifacts like day-of-the-week effects in the series.  Note
that we can equivalently view this (equivalent up to a constant
factor) as predicting the HRR-level COVID-19 case incidence rate \textit{summed}
over some 7-day period in the future, and predicting a binary variable derived
from this.

@@

Recall $Y_{\ell,t}$ denotes the 7-day trailing average of COVID-19 case
incidence rates in location $\ell$ and at time $t$.  Separately for each
$a=7,\ldots,21$, to predict $Y_{\ell,t+a}$ for ahead value $a$, we consider a
simple probabilistic forecasting model of the form:     
\begin{equation}
\label{eq:forecast_ar}
\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s},\ s \leq t)  
= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j}.  
\end{equation}
This model uses current case rates, and the case rates 7 and 14 days ago, in
order to predict (the quantiles of) case rates in the future.  We consider a
total of 7 quantile levels (chosen in accordance with the county-level quantile
levels suggested by the COVID-19 Forecast Hub),      
\begin{equation}
\label{eq:quantile_levels}
\tau \in \{0.025,\ 0.1,\ 0.25,\ 0.5,\ 0.75,\ 0.9,\ 0.975 \}.
\end{equation}
We fit \eqref{eq:forecast_ar} using \textit{quantile regression}
\cite{Koenker:1978, Koenker:2005, Koenker:2006} separately for each $\tau$, 
using data from all 306 HRRs, and within each HRR, using the most recent 21 days
of training data.  This gives us 6,426 training samples for each quantile 
regression problem.

In addition to this pure autoregressive model, we also consider five
probabilistic forecasting models of the form:  
\begin{multline}
\label{eq:forecast_ar_x}
\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s},\ X_{\ell,s},\ s \leq t)  
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
\sum_{j=0}^2 \gamma^{a,\tau}_j X_{\ell,t-7j},
\end{multline}
where $X_{\ell,t}$ denotes any one of the five auxiliary indicators---CHNG-CLI, 
CHNG-COVID, CTIS-CLI-in-community, DV-CLI, or Google-AA---at location $\ell$ and
time $t$. Note that we apply the same lags (current value, along with the values
7 and 14 days ago) for the auxiliary indicators as we do for the case
rates. Training then proceeds just as before: we use the same 7 quantile levels
in \eqref{eq:quantile_levels}, and fit quantile regression separately for each 
level $\tau$, using data from all 306 HRRs and a trailing window of 21 days of
training data.

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{fig/trajectory.pdf}
\caption{This figure displays a forecast from an autoregressive model for the
  HRR containing
  New York City made  
  on October 15 (vertical line).  The fan displays 50\%, 80\% and 95\% intervals
  while the orange curve shows the median forecast. The black curve represents
  ``finalized'' data, as reported in May 2021.} 
\label{fig:trajectory}
\end{figure}

@@

We treat hotspot prediction as a binary classification problem and use a setup 
altogether quite similar to the forecasting setup described previously.
Separately for each $a=7,\ldots,21$, to predict $Z_{\ell,t+a}$, we consider a
simple logistic model:
\begin{equation}
\label{eq:hotspot_ar}
\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s},\ s \leq t) \big) 
= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j},
\end{equation}
where $\logit(p) = \log(p/(1-p))$, the log-odds of $p$.

In addition to this pure autoregressive model, we also consider five
logistic models of the form:  
\begin{multline}
\label{eq:hotspot_ar_x}
\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s},\ X_{\ell,s},\ s \leq t) \big)
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
\sum_{j=0}^2 \gamma^{a,\tau}_j X^\Delta_{\ell,t-7j},
\end{multline}
where we use \smash{$X^\Delta_{\ell,t} = (X_{\ell,t} - X_{\ell,
    t-7})/(X_{\ell,t-7})$}, and again $X_{\ell,t}$ stands for any of the five  
auxiliary indicators at location $\ell$ and time $t$.  We fit the above models,
\eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x}, using logistic regression,   
pooling all 306 HRRs and using a 21-day trailing window for the training data.

@@

For forecasting, we evaluate the probabilistic forecasts produced by the
quantile models in \eqref{eq:forecast_ar} and \eqref{eq:forecast_ar_x} using 
\textit{weighted interval score} (WIS), a quantile-based scoring rule; see e.g.,
\cite{Gneiting:2007}.  WIS is a proper score, meaning that its expectation
is minimized by the population quantiles of the target variable.  The use of WIS
in COVID-19 forecast scoring is discussed in \cite{Bracher:2021}; WIS is also
the main evaluation metric used in the COVID-19 Forecast Hub.

WIS is typically defined for quantile-based forecasts where the quantile levels
are symmetric around 0.5.  This is the case for our choice in
\eqref{eq:quantile_levels}.  Let $F$ be a forecaster comprised of predicted
quantiles $q_\tau$ parametrized by a quantile level $\tau$.  In the case of
symmetric quantile levels, this is equivalent to a collection of central
prediction intervals $(\ell_\alpha, u_\alpha)$, parametrized by an exclusion
probability $\alpha$. The WIS of the forecaster $F$, evaluated at the target
variable $Y$, is defined by:
\begin{equation}
\label{eq:wis_intervals}
\WIS(F,Y) = \sum_\alpha \Big\{\alpha(u_\alpha - \ell_\alpha) + 2 \cdot
\dist(Y, [\ell_\alpha, u_\alpha])\Big\},  
\end{equation}
where $\dist(a, S)$ is the distance between a point $a$ and set $S$ (the
smallest distance between $a$ and an element of $S$).  Note that, corresponding 
to \eqref{eq:quantile_levels}, the exclusion probabilities are $\alpha \in
\{0.05,\ 0.2,\ 0.5,\ 1\}$, resulting in 4 terms in the above sum.  By 
straightforward algebra, it is not hard to see WIS has an alternative 
representation in terms of the predicted quantiles themselves:
\begin{equation}
\label{eq:wis_quantiles}
\WIS(F,Y) = 2 \sum_\tau \phi_\tau(Y - q_\tau), 
\end{equation}
where $\phi_\tau(x) = \tau |x|$ for $x \geq 0$ and $\phi_\tau(x) = (1-\tau)
|x|$ for $x<0$, which is often called the ``tilted absolute'' loss.  While
\eqref{eq:wis_quantiles} is more general (it can accomodate asymmetric quantile
levels), the first form in \eqref{eq:wis_intervals} is typically preferred in
presentation, because the score nicely decouples into a ``sharpness'' component
(first term in each summand) and an ``under/overprediction'' component (second term
in each summand).  The second form given in \eqref{eq:wis_quantiles} is
especially noteworthy in our current study because it reveals WIS to be the same as the
quantile regression loss that we use to train our forecasting models (i.e., our
models are fit to optimize WIS averaged over the training set).

@@

Here, and in the
following subsections, we will use ``AR'' to refer to the pure autoregressive
model both in forecasting, \eqref{eq:forecast_ar}, and in hotspot prediction,
\eqref{eq:hotspot_ar} (the reference to the prediction task should always be
clear from the context). We will also use the name of an auxiliary
indicator---namely ``CHNG-CLI'',  ``CHNG-COVID'', ``CTIS-CLI-in-community'',
``DV-CLI'', or ``Google-AA''---interchangeably with the model in forecasting, 
\eqref{eq:forecast_ar_x}, or hotspot prediction, \eqref{eq:hotspot_ar_x}, that
uses this particular indicator as a feature (the meaning should be clear from
the context).  So, for example, the CHNG-CLI model in forecasting is the one in  
\eqref{eq:forecast_ar_x} that sets $X_{\ell,t}$ to be the value of the CHNG-CLI 
indicator at location $\ell$ and time $t$.
% fits the coefficients based on the training data, and makes its predictions by
% setting $X_{\ell,t}$ to the appropriate CHNG-CLI values.  
Finally, we use the term ``indicator model'' to refer to any one of the ten
models of the form \eqref{eq:forecast_ar_x} or \eqref{eq:hotspot_ar_x} (five from
each of the forecasting and hotspot prediction tasks). We begin with a summary
of the high-level conclusions.

@@

\item In this same analysis, when an indicator model performs better than AR in
  a decreasing period, this tends to co-occur with instances in which the
  indicator ``leads'' case rates (meaning, roughly, on a short-time scale in a
  given location, its behavior mimics that of future case rate behavior).  On
  the other hand, if an an indicator model does better in periods of
  increase, or worse in periods of increase or decrease, its performance is
  unrelated to leadingness.   
\end{itemize}

Finally, to quantify the importance of training and making predictions using
proper vintage data, we ran a parallel set of forecasting and hotspot prediction 
experiments using finalized data. The results, given in the supplement, show
that training and making predictions on finalized data can result in overly
optimistic estimates of true test-time performance (up to 10\% better in terms
of average WIS or AUC). Furthermore, since indicators can have greatly different
backfill profiles, the use of finalized data in retrospective evaluations
changes the relative ranking of models.  For example, CHNG-CLI and DV-CLI,
when trained on finalized data, perform very similarly for the forecasting task.
This makes sense since they are both claims-based indicators supposedly measuring
the same thing.  However, DV-CLI outperforms CHNG-CLI on vintage
data, reflecting its has a less severe backfill profile.

@@

We can also see from the figure that CHNG-COVID and DV-CLI offer the biggest
gains over AR at small ahead values, followed by CTIS-CLI-in-community and
Google-AA, with the former providing the biggest gains at large ahead values.
The CHNG-CLI model performs basically the same as AR.  This is likely due to the
fact that CHNG-CLI suffers from spatially heterogeneous volatility due to
backfill.  The evidence for this explanation is twofold: (1) the CHNG-CLI model benefits from a
more robust method of aggregating WIS (geometric mean; shown in the supplement);
and (2) when we train and make predictions on finalized data, it handily beats
AR, on par with the best-performing models (also shown in the 
supplement).
% TODO note what we said here about the supplement

@@

To test this hypothesis, we reran all of the prediction experiments but with
$X_{\ell,t}$ in the each indicator models
%in \eqref{eq:forecast_ar_x} \eqref{eq:hotspot_ar_x}
replaced with suitable random noise (bootstrap samples
from a signal's historical distribution).  The results, shown and explained more
precisely in the supplement, are vastly different (worse) than the original set
of results.  In both forecasting and hotspot prediction, the ``fake'' indicator
models
% (with random noise as auxiliary features)
offered essentially no
improvement over the pure AR model, which provides a strong rejection
(informally speaking) of the implicit regularization hypothesis.

On the topic of regularization, it is also worth noting that the use of 
$\ell_1$ regularization (tuned by cross-validation) in fitting any of the models
in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar}, or 
\eqref{eq:hotspot_ar_x} did not generally improve their performance
(experiments not shown).  This is likely due to the fact that the number of
training samples is large compared to the number of features (6,426 training
samples and only 3--6 features). [DJM: and that proper CV requires longer
training windows, exacerbating the nonstationarity, a scenario in which its
behavior is poorly understood.]

% \begin{figure}[t]
% \includegraphics[width=\columnwidth]{fig/hot-1.pdf}
% \caption{Area under the curve for each hotspot prediction model, calculated over
%   all prediction dates and all HRRs.} 
% \label{fig:hotspot}
% \end{figure}

@@

% Essentially, during downswings, improved performance is strongly correlated
% with predicing larger declines than the AR model. The AR model tends to
% overpredict relative to the decline, but adding indicators tends to diminish
% this effect. During upswings, the AR model tends to underpredict, but when the
% addition of an indicator improves performance (a more rare occurrence) it is 
% rarely due to higher median predictions.

% RJT: Commented out the last paragraph for now, because this is kind of a tough
% thing to explain succintly. Can revisit if we really want to include it here,
% or just let them find it in the supplement?
% DJM: Agreed. In the supplement.

@@

\acknow{We thank Matthew Biggerstaff, Logan Brooks, Johannes Bracher, Michael 
  Johansson, Evan Ray, Nicholas Reich, and Roni Rosenfeld for several
  enlightening conversations about forecast evaluation. The research was
  supported by Centers for Disease Control and Prevention (CDC) grant
  U01IP001121; gifts from Facebook, Google, McCune, and Optum; the National
  Science Foundation Graduate Research Fellowship Program (NSF GRFP) award
  DGE175016; the Canadian Statistical Sciences Institute; and the National
  Sciences and Engineering Research Council of Canada (NSERC) RGPIN-2021-02618.}

>>>>

% Please add a significance statement to explain the relevance of your work
\significancestatement{Forecasting is a critical element in the public health
  response to any fast-moving epidemic or pandemic.  Predicting the future
  spread of some disease process, based on disease activity in the recent
  past---say, using what is known as autoregressive modeling in the time series
  literature---is a basic and widely-used method.  In this paper we study
  whether including auxiliary indicators as additional features in such an
  autoregressive model can be useful in COVID-19 case forecasting and hotspot
  prediction.  We find that indicators based on de-identified medical insurance 
  claims, self-reported symptoms through online surveys, and COVID-related
  internet search activity all provide a nontrivial boost in accuracy, though
  their utility generally depends the pandemic's local dynamics (upswing,
  downswing, or flat period) at prediction time.
}

@@

We consider prediction of future COVID-19 case rates or case hotspots (to be
defined precisely shortly).  By case rate, we mean the case count per 100,000
people (the standard in epidemiology).  We use reported case data aggregated by 
JHU CSSE \cite{Dong:2020}, which, like the auxiliary indicators that we use to
supplement the basic autoregressive models, is accessible through the COVIDcast
API \cite{CovidcastAPI}.

The indicators we focus on provide information not generally available from
standard public health reporting. Among the many auxiliary indicators collected
in the API, we study the following five:  
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Trends and Impact Survey COVID-like illness in the community
  (CTIS-CLI-in-community): The estimated percentage of the population who know
  someone in their local community that is sick, based on Delphi's COVID Trends
  and Impact Survey, in partnership with Facebook.
\item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
  computed based on de-identified medical insurance claims from other health
  systems partners.  
\item Google Search Trends for anosmia and ageusia (Google-AA): A measure of 
  Google search volume for queries that relate to anosmia or ageusia (loss of
  smell or taste), based on Google's Search Trends data set.  
\end{itemize}
In short, we choose these indicators because, conceptually speaking, they
measure aspects of an individual's disease progression that would plausibly
precede the occurence of (at worst, co-occur  with) the report of a positive
COVID-19 test, through standard public health reporting streams.

@@

HRRs, by their definition, would be most relevant for forecasting hospital
demand.  We have chosen to focus on cases (forecasting and predicting
hotspots) at the HRR level because the indicators considered should be more 
useful in predicting case activity rather than hospital demand, as the former is  
intuitively more contemporaneous to the events that are measured by
the given five indicators. Predicting case rates (and hotspots) at the HRR level
is still a reasonable goal in its own right; and moreover, it could be used to
feed predicted case information into downstream hospitalization models.

@@

In this paper, all models are fit with ``vintage'' training data. This means
that for a given prediction date, say, September 28, 2020, we train models 
using data that would have been available to us \textit{as of}
September 28 (imagine that we can ``rewind'' the clock to September 28 and query
the COVIDcast API to get the latest data it would have had available at that
point in time.)  This is possible because of the COVIDcast API's comprehensive
data versioning system (described in more detail in our companion paper).  We
also use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the 
process of training arbitrary prediction models over a sequence of prediction
dates, by constructing the proper sequence of vintage training data sets.

\begin{figure}[tb!]
  \includegraphics[width=\columnwidth]{fig/revisions.pdf}
  % RJT: Ensuring we start with a "declarative title" :) 
  % https://www.internationalscienceediting.com/how-to-write-a-figure-caption/
  \caption{Revision behavior for two signals in the HRR containing New York  
    City.  Each colored line corresponds to the data as reported on a particular 
    date (\textit{as of} dates varying from September 28 through October
    19). The left panel shows the DV-CLI signal, which was regularly revised
    throughout the period, though the effects fade as we look further back in   
    time. In contrast, the right panel shows case rates reported by JHU CSSE,
    which remain ``as reported'' on September 28, with a huge spike towards the 
    end of this period, until a major correction is made on October 19, which
    brings down this spike and affects all prior data as well.}
  \label{fig:vintage}
\end{figure}

Vintage training data means different things, in practice, for different
signals. The three signals based on medical claims, CHNG-CLI, CHNG-COVID, and
DV-CLI, are most often 3-5 days latent, and subject to a considerable but
regular degree of revision or ``backfill'' after their initial publication date.
The survey-based signal, CTIS-CLI-in-community, is 2 days latent, and rarely
undergoes any revision at all.  The target variable itself, reported COVID-19
case rates, is 1 day latent, and exhibits frequent, unpredictable revisions
after intial publication.  Compared to the revision pattern in the medical
claims signals, which are much more systematic in nature, revisions in case
reports can be highly erratic. Big spikes or other anomalies can occur in the
data as reporting backlogs are cleared, changes in case definitions are made,
etc. Groups like JHU CSSE then work tirelessly to correct such anomalies after
first publication (e.g., they will attempt to back-distribute a spike when a
reporting backlog is cleared, by working with a local authority to figure out
how this should best be done), which can result in very nontrivial revisions.
See Figure~\ref{fig:vintage} for an example.

@@

Why do we define the response variables via 7-day averaging? The short answer 
is robustness: averaging stabilizes the case time series, and accounts for 
uninteresting artifacts like day-of-the-week effects in the series.  Note that
we can also equivalently view this (equivalent up to a constant factor) as
predicting the HRR-level case incidence rate \textit{summed} over some 7-day
period in the future, and predicting a binary variable derived from this.

@@

Recall $Y_{\ell,t}$ denotes the 7-day trailing average of COVID-19 case
incidence rates in location $\ell$ and at time $t$.  Separately for each
$a=7,\ldots,21$, to predict $Y_{\ell,t+a}$ for ahead value $a$, we consider a
simple probabilistic forecasting model of the form:     
\begin{equation}
\label{eq:forecast_ar}
\Quantile_\tau(Y_{\ell,t+a} \ | \ Y_{\ell,s},\ s \leq t)   
= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j}.  
\end{equation}
This model uses current case rates, and the case rates 7 and 14 days ago, in
order to predict (the quantiles of) case rates in the future.  We consider a
total of 7 quantile levels (chosen in accordance with the county-level quantile
levels suggested by the COVID-19 Forecast Hub),      
\begin{equation}
\label{eq:quantile_levels}
\tau \in \{0.025,\ 0.1,\ 0.25,\ 0.5,\ 0.75,\ 0.9,\ 0.975 \}.
\end{equation}
We fit \eqref{eq:forecast_ar} using \textit{quantile regression}
\cite{Koenker:1978, Koenker:2005, Koenker:2006} separately for each $\tau$, 
using data from all 306 HRRs, and within each HRR, using the most recent 21 days
of training data.  This gives us 6,426 training samples for each quantile 
regression problem.

In addition to this pure autoregressive model, we also consider five
probabilistic forecasting models of the form:  
\begin{multline}
\label{eq:forecast_ar_x}
\Quantile_\tau(Y_{\ell,t+a} \ | \ Y_{\ell,s},\ X_{\ell,s},\ s \leq t) 
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
\sum_{j=0}^2 \gamma^{a,\tau}_j X_{\ell,t-7j},
\end{multline}
where $X_{\ell,t}$ denotes any one of the five auxiliary indicators---CHNG-CLI, 
CHNG-COVID, CTIS-CLI-in-community, DV-CLI, or Google-AA---at location $\ell$ and
time $t$. Note that we apply the same lags (current value, along with the values
7 and 14 days ago) for the auxiliary indicators as we do for the case
rates. Training then proceeds just as before: we use the same 7 quantile levels
in \eqref{eq:quantile_levels}, and fit quantile regression separately for each 
level $\tau$, using data from all 306 HRRs and a trailing window of 21 days of
training data.

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{fig/trajectory.pdf}
\caption{Forecast for the HRR containing New York City from an autoregressive
  model made on October 15 (vertical line).  The fan displays 50\%, 80\% and
  95\% intervals while the orange curve shows the median forecast. The black
  curve shows ``finalized'' data, as reported in May 2021.}   
\label{fig:trajectory}
\end{figure}

@@

We treat hotspot prediction as a binary classification problem and use a setup 
altogether quite similar to the forecasting setup described previously.
Separately for each $a=7,\ldots,21$, to predict $Z_{\ell,t+a}$, we consider a
simple logistic model:
\begin{equation}
\label{eq:hotspot_ar}
\logit \big( \P(Z_{\ell,t+a} = 1 \ | \ Y_{\ell,s},\ s \leq t) \big) 
= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j},
\end{equation}
where $\logit(p) = \log(p/(1-p))$, the log-odds of $p$.

In addition to this pure autoregressive model, we also consider five
logistic models of the form:  
\begin{multline}
\label{eq:hotspot_ar_x}
\logit \big( \P(Z_{\ell,t+a} = 1 \ | \ Y_{\ell,s},\ X_{\ell,s},\ s \leq t) \big)
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
\sum_{j=0}^2 \gamma^{a,\tau}_j X^\Delta_{\ell,t-7j},
\end{multline}
where we use \smash{$X^\Delta_{\ell,t} = (X_{\ell,t} - X_{\ell,
    t-7})/(X_{\ell,t-7})$}, and again $X_{\ell,t}$ stands for any of the five  
auxiliary indicators at location $\ell$ and time $t$.  We fit the above models,
\eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x}, using logistic regression,   
pooling all 306 HRRs and using a 21-day trailing window for the training data.

@@

For forecasting, we evaluate the probabilistic forecasts produced by the
quantile models in \eqref{eq:forecast_ar} and \eqref{eq:forecast_ar_x} using 
\textit{weighted interval score} (WIS), a quantile-based scoring rule; see e.g.,
\cite{Gneiting:2007}.  WIS is a proper score, which means that its expectation 
is minimized by the population quantiles of the target variable.  The use of WIS
in COVID-19 forecast scoring is discussed in \cite{Bracher:2021}; WIS is also
the main evaluation metric used in the COVID-19 Forecast Hub.

WIS is typically defined for quantile-based forecasts where the quantile levels
are symmetric around 0.5.  This is the case for our choice in
\eqref{eq:quantile_levels}.  Let $F$ be a forecaster comprised of predicted
quantiles $q_\tau$ parametrized by a quantile level $\tau$.  In the case of
symmetric quantile levels, this is equivalent to a collection of central
prediction intervals $(\ell_\alpha, u_\alpha)$, parametrized by an exclusion
probability $\alpha$. The WIS of the forecaster $F$, evaluated at the target
variable $Y$, is defined by:
\begin{equation}
\label{eq:wis_intervals}
\WIS(F,Y) = \sum_\alpha \Big\{\alpha(u_\alpha - \ell_\alpha) + 2 \cdot
\dist(Y, [\ell_\alpha, u_\alpha])\Big\},  
\end{equation}
where $\dist(a, S)$ is the distance between a point $a$ and set $S$ (the
smallest distance between $a$ and an element of $S$).  Note that, corresponding  
to \eqref{eq:quantile_levels}, the exclusion probabilities are $\alpha \in
\{0.05,\ 0.2,\ 0.5,\ 1\}$, resulting in 4 terms in the above sum.  By 
straightforward algebra, it is not hard to see WIS has an alternative 
representation in terms of the predicted quantiles themselves:
\begin{equation}
\label{eq:wis_quantiles}
\WIS(F,Y) = 2 \sum_\tau \phi_\tau(Y - q_\tau), 
\end{equation}
where $\phi_\tau(x) = \tau |x|$ for $x \geq 0$ and $\phi_\tau(x) = (1-\tau)
|x|$ for $x<0$, which is often called the ``tilted absolute'' loss.  While
\eqref{eq:wis_quantiles} is more general (it can accomodate asymmetric quantile
levels), the first form in \eqref{eq:wis_intervals} is typically preferred in
presentation, as the score nicely decouples into a ``sharpness'' component
(first term in each summand) and an ``under/overprediction'' component (second
term in each summand).  But the second form given in \eqref{eq:wis_quantiles} is 
especially noteworthy in our current study because it reveals WIS to be the same
as the quantile regression loss that we use to train our forecasting models
(i.e., we fit by optimizing WIS averaged over the training data).

@@

Here, and in the following subsections, we will use ``AR'' to refer to the pure
autoregressive model both in forecasting, \eqref{eq:forecast_ar}, and in hotspot
prediction, \eqref{eq:hotspot_ar} (the reference to the prediction task should
always be clear from the context). We will also use the name of an auxiliary 
indicator---namely ``CHNG-CLI'',  ``CHNG-COVID'', ``CTIS-CLI-in-community'',
``DV-CLI'', or ``Google-AA''---interchangeably with the model in forecasting, 
\eqref{eq:forecast_ar_x}, or hotspot prediction, \eqref{eq:hotspot_ar_x}, that
uses this particular indicator as a feature (the meaning should be clear from
the context).  So, for example, the CHNG-CLI model in forecasting is the one in  
\eqref{eq:forecast_ar_x} that sets $X_{\ell,t}$ to be the value of the CHNG-CLI 
indicator at location $\ell$ and time $t$.
% fits the coefficients based on the training data, and makes its predictions by
% setting $X_{\ell,t}$ to the appropriate CHNG-CLI values.  
Finally, we use the term ``indicator model'' to refer to any one of the ten
models of the form \eqref{eq:forecast_ar_x} or \eqref{eq:hotspot_ar_x} (five
from each of the forecasting and hotspot prediction tasks).

We begin with a summary of the high-level conclusions.

@@

\item In this same analysis, when an indicator model performs better than AR in
  a decreasing period, this tends to co-occur with instances in which the
  indicator ``leads'' case rates (meaning, roughly, on a short-time scale in a
  given location, its behavior mimics that of future case rates).  On
  the other hand, if an an indicator model does better in periods of
  increase, or worse in periods of increase or decrease, its performance is
  not as related to leadingness.   
\end{itemize}

Finally, to quantify the importance of training and making predictions using
proper vintage data, we ran a parallel set of forecasting and hotspot prediction 
experiments using finalized data. The results, given in the supplement, show
that training and making predictions on finalized data can result in overly
optimistic estimates of true test-time performance (up to 10\% better in terms
of average WIS or AUC). Furthermore, since indicators can have greatly different
backfill profiles, the use of finalized data in retrospective evaluations
changes the relative ranking of models.  For example, CHNG-CLI and DV-CLI,
when trained on finalized data, perform very similarly in forecasting.
This makes sense since they are both claims-based indicators that are supposedly
measuring the same thing.  However, DV-CLI outperforms CHNG-CLI on vintage
data, reflecting its has a less severe backfill profile.

@@

We can also see from the figure that CHNG-COVID and DV-CLI offer the biggest
gains over AR at small ahead values, followed by CTIS-CLI-in-community and
Google-AA, with the former providing the biggest gains at large ahead values.
The CHNG-CLI model performs basically the same as AR.  This is likely due to the
fact that CHNG-CLI suffers from volatility due to backfill.  The evidence for
this explanation is twofold: (1) the CHNG-CLI model benefits from a more robust
method of aggregating WIS (geometric mean; shown in the supplement); and (2)
when we train and make predictions on finalized data, it handily beats AR, on
par with the best-performing models (also shown in the supplement).

@@

To test this hypothesis, we reran all of the prediction experiments but with
$X_{\ell,t}$ in the each indicator models replaced with suitable random noise
(bootstrap samples from a signal's history).  The results, shown and explained
more precisely in the supplement, are vastly different (worse) than the original
set of results.  In both forecasting and hotspot prediction, the ``fake''
indicator models offered essentially no improvement over the pure AR model,
which strongly rejects (informally speaking) the implicit regularization
hypothesis.

On the topic of regularization, it is also worth noting that the use of 
$\ell_1$ regularization (tuned by cross-validation) in fitting any of the models
in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar}, or 
\eqref{eq:hotspot_ar_x} did not generally improve their performance
(experiments not shown).  This is likely due to the fact that the number of
training samples is large compared to the number of features (6,426 training
samples and only 3--6 features). 
% DJM: and that proper CV requires longer training windows, exacerbating the
% nonstationarity, a scenario in which its behavior is poorly understood.

@@

@@

\acknow{We thank Matthew Biggerstaff, Logan Brooks, Johannes Bracher, Michael 
  Johansson, Evan Ray, Nicholas Reich, and Roni Rosenfeld for several
  enlightening conversations about forecast evaluation. The research was
  supported by Centers for Disease Control and Prevention (CDC) grant
  U01IP001121; gifts from Facebook, Google, McCune, and Optum; the Canadian
  Statistical Sciences Institute; National Sciences and Engineering Research
  Council of Canada (NSERC) RGPIN-2021-02618; and National Science Foundation
  Graduate Research Fellowship Program (NSF GRFP) award DGE175016.}
